Artin Group Presentations Arising from Cluster Algebras

### 2016-10-21 Paper.tex
Corrections

In \cite{FZ02}, Fomin and Zelevinsky first introduced the concept of cluster algebras. Barot and Marsh extended Fomin and Zelevinsky's results in \cite{BM13}, providing a presentation of the reflection group associated to a Dynkin diagram with generators that correspond to elements of a companion basis associated to a seed of a finite type cluster algebra. They define generator relations corresponding to chordless cycles arising in diagrams of finite type in order to give a Coxeter group presentation for these diagrams. They also proved that this Coxeter group presentation is invariant up to isomorphism under the mutation equivalence relation. That is, given a diagram $\Gamma$ and a diagram mutation equivalent to $\Gamma$, denoted $\Gamma^{\prime} = \mu_k(\Gamma)$, they proved that $W_{\Gamma} \cong W_{\Gamma^{\prime}}$, where $W_{\Gamma}$ and $W_{\Gamma^{\prime}}$ are the group presentations corresponding to $\Gamma$ and $\Gamma^{\prime}$, respectively.

@@

A \textit{cluster algebra} is an integral domain which can be generated by a set of elements called \textit{cluster variables} that satisfy certain exchange relations. Following the style of \cite{FZ02} and \cite{BM13}, we will define cluster algebras in terms of \textit{skew-symmetrisable} matrices (that is, a matrix $B$ such that there exists a diagonal matrix $D$ of the same size with $D_{ii} >0$ such that $DB$ is skew-symmetric). Let $\mathbb{F} = \mathbb{Q}(u_1, u_2, \ldots, u_n)$ be the field of rational functions in $n$ indeterminates over $\mathbb{Q}$. We will define an \textit{initial seed} for the cluster algebra to be a fixed pair $(\textbf{x}, B)$, where $\textbf{x} = \{x_1, \ldots, x_n\}$ is a free generating set of $\mathbb{F}$ and $B$ is an $n \times n$ skew-symmetrisable matrix. Define $x_k^{\prime} \in \mathbb{F}$ by the \textit{exchange relation}
\begin{displaymath}
x_k^{\prime}x_k = \prod_{B_{ik} > 0}{x_i^{B_{ik}}} + \prod_{B_{ik} < 0}{x_i^{-B_{ik}}}
\end{displaymath}
Then, given an initial seed $(\textbf{x}, B)$ and $k \in {1,2,\ldots,n}$, we can define a \textit{mutation} of the seed at $k$, denoted $\mu_k(\boldsymbol{x}, B) = (\textbf{x}^{\prime}, B^{\prime})$ where:
\begin{displaymath}
B_{ij}^{\prime} = \begin{cases} - B_{ij} & \mbox{ if } i = k \mbox{ or } j = k;\\
B_{ij} + \frac{|B_{ik}|B_{kj} + B_{ik}|B_{kj}|}{2} & \mbox{ otherwise. }\\
\end{cases}
\end{displaymath}
and $\textbf{x}^{\prime} = \left\{ x_1, x_2, \ldots, x_{k-1}, x_k^{\prime}, x_{k+1}, \ldots, x_n \right\}$.
Such a mutation or a sequence of such mutations generate \textit{seeds} which in turn generate all cluster variables in that, for each $\textbf{x} = \left\{ x_1, \ldots, x_n \right\}$ corresponding to a seed of the cluster algebra, the entries $x_i$ are the cluster variables. \\

@@

\begin{rem}
We purposely include relations $(T2),(T3)$ but not $(T1)$ in order to make our relation labeling analogous to that of \cite{BM13} at the beginning of Section 3. Note that if we add the additional relation $(R1)$ as defined at the beginning of Section 3 of \cite{BM13} (namely, if we add $s_i^2 = e$ for all vertices $i$ in $\Gamma$), then we will precisely obtain the group $W_\Gamma$ as defined at the beginning of Section 3 in \cite{BM13}.
\end{rem}

@@

\begin{ex}
The relations $(T2),(T3)$ in that $\Gamma$ is a triangle with two edges of weight 2 and one of weight 1 are as follows:
\[\Gamma = \begin{tikzpicture}[baseline=-0.5ex]
\matrix[matrix of math nodes,column sep={25pt,between origins},row
    sep={40pt,between origins}] at (0,0)
  {
    && |[name = u]|\circ& \\
    %
    &|[name=dl]| \circ & &|[name=dr]| \circ \\
  };
\draw[->]
   (u) to node[right] {2}(dr)
  ;
  \draw[->]
  (dl) to node[left]{2}(u)
  ;
  \draw[->]
  (dr) to node[below]{1}(dl)
  ;
  \node [yshift = .4 cm] at (u){1};
    \node [yshift = -.4 cm,xshift = .4 cm] at (dr){2};
    \node [yshift = -.4 cm,xshift = -.4 cm] at (dl){3};
   \end{tikzpicture}\]
   \begin{enumerate}
\item[(T2)]
\begin{itemize}
\item $\langle s_1,s_2 \rangle^4 = \langle s_2,s_1 \rangle^4,$ i.e., $s_1s_2s_1s_2 = s_2s_1s_2s_1$
\item $\langle s_2,s_3 \rangle^3 = \langle s_3,s_2 \rangle^3,$ i.e., $s_2s_3s_2 = s_3s_2s_3$
\item $\langle s_3,s_4 \rangle^4 = \langle s_4,s_3 \rangle^4,$ i.e., $s_3s_4s_3s_4 = s_4s_3s_4s_3$              
\end{itemize}
\item[(T3)]
\begin{itemize}
\item $s_1s_2^{-1}s_3^{-1}s_2s_1^{-1}s_2^{-1}s_3s_2 = e$

\item $s_2s_3^{-1}s_1^{-1}s_3s_2^{-1}s_3^{-1}s_1s_3 =e$

@@

\begin{proof}
The lemma follows from the fact that
\begin{align*}
& s_{k}^{-1}s_{j}(s_{i}p(i, j)s_{i}^{-1}p(i, j)^{-1})s_{j}^{-1}s_{k} \\
&= s_{k}^{-1}s_{j}(s_{i}s_{j}^{-1}s_{k}s_{j}s_{i}^{-1}s_{j}^{-1}s_{k}^{-1}s_{j})s_{j}^{-1}s_{k} \\
&= s_{k}^{-1}s_{j}s_{i}s_{j}^{-1}s_{k}s_{j}s_{i}^{-1}s_{j}^{-1} \\
&= s_{k}^{-1}s_{i}^{-1}s_{j}s_{i}s_{k}s_{i}^{-1}s_{j}^{-1}s_{i} \\
&= s_{k}^{-1}p(k, i)s_{k}p(k, i)^{-1}
\end{align*}
\end{proof}

@@

@@

\bibitem[BM13]{BM13}
Michael Barot and Robert~J. Marsh.
\newblock Reflection group presentations arising from cluster algebras.
\newblock arXiv: 1112.2300v2, 2013.

\bibitem[Cha06]{C06}
Ruth Charney.
\newblock An introduction to right-angled artin groups.
\newblock arXiv:math/0610668, 2006.

@@

\bibitem[FZ02]{FZ02}
Sergey Fomin and Andrei Zelevinsky.
\newblock Cluster algebras I: Foundations.
\newblock {\em J. Amer. Math. Soc.}, 15:497--529, 2002.

\bibitem[FZ03]{FZ03}
Sergey Fomin and Andrei Zelevinsky.
\newblock Cluster algebras II: Finite type classification.
\newblock {\em Invent. Math.}, 154:63--121, 2003.

\bibitem[GM14]{GM14}
Joseph Grant and Robert~J. Marsh.
\newblock Braid groups and quiver mutation.
\newblock arXiv:1408.5267, 2014.

>>>>

In \cite{FZ02}, Fomin and Zelevinsky first introduced the concept of cluster algebras. Barot and Marsh extended Fomin and Zelevinsky's results in \cite{BM13}, providing a presentation of the reflection group associated to a Dynkin diagram with generators that correspond to elements of a companion basis associated to a seed of a finite type cluster algebra. They define generator relations corresponding to chordless cycles arising in diagrams of finite type in order to give a Coxeter group presentation for these diagrams. They also proved that this Coxeter group presentation is invariant up to isomorphism under the mutation equivalence relation. That is, given a diagram $\Gamma$ and a diagram $\Gamma^{\prime}$ which is mutation equivalent to $\Gamma$, they proved that $W_{\Gamma} \cong W_{\Gamma^{\prime}}$, where $W_{\Gamma}$ and $W_{\Gamma^{\prime}}$ are the group presentations corresponding to $\Gamma$ and $\Gamma^{\prime}$, respectively.

@@

A \textit{cluster algebra} is an integral domain which can be generated by a set of elements called \textit{cluster variables} that satisfy certain exchange relations. Following the style of \cite{FZ02} and \cite{BM13}, we will define cluster algebras in terms of \textit{skew-symmetrisable} matrices (that is, a matrix $B$ such that there exists a diagonal matrix $D$ of the same size with $D_{ii} >0$ such that $DB$ is skew-symmetric). Let $\mathbb{F} = \mathbb{Q}(u_1, u_2, \ldots, u_n)$ be the field of rational functions in $n$ indeterminates over $\mathbb{Q}$. We will define an \textit{initial seed} for the cluster algebra to be a fixed pair $(\textbf{x}, B)$, where $\textbf{x} = \{x_1, \ldots, x_n\}$ is a free generating set of $\mathbb{F}$ and $B$ is an $n \times n$ skew-symmetrisable matrix. Define $x_k^{\prime} \in \mathbb{F}$ by the \textit{exchange relation}
\begin{displaymath}
x_k^{\prime}x_k = \prod_{B_{ik} > 0}{x_i^{B_{ik}}} + \prod_{B_{ik} < 0}{x_i^{-B_{ik}}}
\end{displaymath}
Then, given an initial seed $(\textbf{x}, B)$ and $k \in {1,2,\ldots,n}$, we can define a \textit{mutation} of the seed at $k$, denoted $\mu_k(\textbf{x}, B) = (\textbf{x}^{\prime}, B^{\prime})$ where:
\begin{displaymath}
B_{ij}^{\prime} = \begin{cases} - B_{ij} & \mbox{ if } i = k \mbox{ or } j = k;\\
B_{ij} + \frac{|B_{ik}|B_{kj} + B_{ik}|B_{kj}|}{2} & \mbox{ otherwise. }\\
\end{cases}
\end{displaymath}
and $\textbf{x}^{\prime} = \left\{ x_1, x_2, \ldots, x_{k-1}, x_k^{\prime}, x_{k+1}, \ldots, x_n \right\}$.
Such a mutation or a sequence of such mutations generate \textit{seeds} which in turn generate all cluster variables in that, for each $\textbf{x} = \left\{ x_1, \ldots, x_n \right\}$ corresponding to a seed of the cluster algebra, the entries $x_i$ are the cluster variables. \\

@@

\begin{rem}
We purposely include relations $(T2),(T3)$ but not $(T1)$ in order to make our relation labeling analogous to that of \cite{BM13} at the beginning of Section 3. Note that if we add the additional relation $(R1)$ as defined at the beginning of Section 3 of \cite{BM13} (namely, if we add $s_i^2 = e$ for all vertices $i$ in $\Gamma$), then we will precisely obtain the group $W_\Gamma$ as defined at the beginning of Section 3 in \cite{BM13}. (It is clear that we obtain the relations of the form $(R2)$ in $W_{\Gamma}$ from the $(T2)$ relations in $A_{\Gamma}$, and we obtain the relations of the form $(R3)(a)$ and $(R3)(b)$ for $k = 1$ in $W_{\Gamma}$ from the $(T3)$ relations in $A_{\Gamma}$. From here, we obtain the remaining $(R3)(b)$ relations in $W_{\Gamma}$ by Lemmas 4.2 and 4.3 in \cite{BM13}.  
\end{rem}

@@

\begin{ex}
The relations $(T2),(T3)$ in that $\Gamma$ is a triangle with two edges of weight 2 and one of weight 1 are as follows:
\[\Gamma = \begin{tikzpicture}[baseline=-0.5ex]
\matrix[matrix of math nodes,column sep={25pt,between origins},row
    sep={40pt,between origins}] at (0,0)
  {
    && |[name = u]|\circ& \\
    %
    &|[name=dl]| \circ & &|[name=dr]| \circ \\
  };
\draw[->]
   (u) to node[right] {2}(dr)
  ;
  \draw[->]
  (dl) to node[left]{2}(u)
  ;
  \draw[->]
  (dr) to node[below]{1}(dl)
  ;
  \node [yshift = .4 cm] at (u){1};
    \node [yshift = -.4 cm,xshift = .4 cm] at (dr){2};
    \node [yshift = -.4 cm,xshift = -.4 cm] at (dl){3};
   \end{tikzpicture}\]
   \begin{enumerate}
\item[(T2)]
\begin{itemize}
\item $\langle s_1,s_2 \rangle^4 = \langle s_2,s_1 \rangle^4,$ i.e., $s_1s_2s_1s_2 = s_2s_1s_2s_1$
\item $\langle s_2,s_3 \rangle^3 = \langle s_3,s_2 \rangle^3,$ i.e., $s_2s_3s_2 = s_3s_2s_3$
\item $\langle s_3,s_1 \rangle^4 = \langle s_1,s_3 \rangle^4,$ i.e., $s_3s_1s_3s_1 = s_1s_3s_1s_3$              
\end{itemize}
\item[(T3)]
\begin{itemize}
\item $s_1s_2^{-1}s_3s_2s_1^{-1}s_2^{-1}s_3^{-1}s_2 = e$

\item $s_2s_3^{-1}s_1s_3s_2^{-1}s_3^{-1}s_1^{-1}s_3 =e$

@@

\begin{proof}
The lemma follows from the fact that
\begin{align*}
& s_{k}^{-1}s_{j}(s_{i}p(i, j)s_{i}^{-1}p(i, j)^{-1})s_{j}^{-1}s_{k} \\
&= s_{k}^{-1}s_{j}(s_{i}s_{j}^{-1}s_{k}s_{j}s_{i}^{-1}s_{j}^{-1}s_{k}^{-1}s_{j})s_{j}^{-1}s_{k} \\
&= s_{k}^{-1}s_{j}s_{i}s_{j}^{-1}s_{k}s_{j}s_{i}^{-1}s_{j}^{-1} \\
&= s_{k}^{-1}s_{i}^{-1}s_{j}s_{i}s_{k}s_{i}^{-1}s_{j}^{-1}s_{i} \\
&= s_{k}^{-1}p(k, i)s_{k}p(k, i)^{-1},
\end{align*}
and from the fact that $s_{k}^{1}p(k,i)s_{k}p(k,i)^{-1} = e$ if and only if $$p(k,i)s_{k} = s_k(s_{k}^{-1}p(k,i)s_{k}p(k,i)^{-1})p(k,i) = s_kp(k,i)$$.
\end{proof}

@@

Carrying out computations similar to those in Section 6, one can check directly that with this choice of relations, we have that for each diagram $\Gamma$ occuring in the first column of Table \ref{table2} and for each vertex $k$ of $\Gamma$, the map $\phi: A_{\mu_k(\Gamma)} \rightarrow A_{\Gamma}$ defined as in Proposition 6.2 is an isomorphism. Thus we conjecture the following:

@@

\bibitem[BM13]{BM13}
M.~Barot and R.~Marsh.
\newblock Reflection group presentations arising from cluster algebras.
\newblock arXiv: 1112.2300v2, 2013.

\bibitem[Cha06]{C06}
R.~Charney.
\newblock An introduction to right-angled artin groups.
\newblock arXiv:math/0610668, 2006.

@@

\bibitem[FZ02]{FZ02}
S.~Fomin and A.~Zelevinsky.
\newblock Cluster algebras I: Foundations.
\newblock {\em J. Amer. Math. Soc.}, 15:497--529, 2002.

\bibitem[FZ03]{FZ03}
S.~Fomin and A.~Zelevinsky.
\newblock Cluster algebras II: Finite type classification.
\newblock {\em Invent. Math.}, 154:63--121, 2003.

\bibitem[GM14]{GM14}
J.~Grant and R.~Marsh.
\newblock Braid groups and quiver mutation.
\newblock arXiv:1408.5267, 2014.
